---
title: "South Africa’s biodiversity profile"
---

```{r, echo=FALSE}

library(gt)

```

------------------------------------------------------------------------

# Overview of South Africa's biodiversity profile

South Africa has exceptional biodiversity, characterised by high species richness, high levels of species endemism and a wide variety of ecosystems. South Africa's diversity and richness are not limited to biodiversity. Within its borders are also diverse cultures and languages, and exceptional geological and climatic diversity.

Identified as one of the world's 17 megadiverse nations, South Africa ranks as one of the top ten nations globally for plant species richness and third for marine species endemism. With a landmass of 1.2 million km2 and surrounding seas of 1.1 million km2, South Africa is among the smaller of the world's megadiverse countries -- which together contain more than two-thirds of the world's biodiversity. South Africa also holds three of the world's 35 biodiversity hotspots (a measure of biological diversity combined with vulnerability to threats): the Cape Floristic Region, Succulent Karoo biome, and the Maputaland--Pondoland--Albany centre of endemism.

South Africa's wide range of bioclimatic, oceanographic, geological and topographical settings have resulted not only in high species diversity and endemism, but also high ecosystem diversity and endemism across all realms. There is a wide variety of terrestrial biomes and marine ecoregions in South Africa, its surrounding seas and sub-Antarctic territory; ranging from the unique Fynbos biome to the extensive savannas and grasslands of the eastern interior, and from the Subtropical Indian Ocean through the Warm Temperate Agulhas Shelf to the cold upwelling influenced shelf of the Southern Benguela (@tbl-speciesrichness). Situated 1 700 km south of the country, the Prince Edward Islands and their surrounding seas add a cold, sub-Antarctic set of ecoregions and biomes to South Africa's territory (@tbl-speciesrichness).

```{r, echo=FALSE}
#| label: tbl-speciesrichness
#| tbl-cap: Summary of species richness and endemism for selected South African taxonomic groups, including global estimates.
#| echo: false
#| message: false
#| warning: false

Dataframe3 <- matrix(nrow = 8, ncol = 6)
colnames(Dataframe3) <- c("Taxonomic group", "Species", "Endemics", "Endemics as % of total SA species", "Global estimates of number of species", "Species in SA as % of global total")

Dataframe3[1,] <- c("Amphibians", "125", "62", "50%", "7 934", "1.6%")

Dataframe3[2,] <- c("Birds", "732", "38", "5%", "11 122", "6.6%")
Dataframe3[3,] <- c("Butterflies", "799", "418", "52%", "17 500", "4.5%")
Dataframe3[4,] <- c("Dragonflies", "162","28", "17%", "5 680", "2.9%")
Dataframe3[5,] <- c("Fishes (freshwater)", "118", "58", "49%", "14 953", "0.8%")
Dataframe3[6,] <- c("Fishes (marine)", "~ 2 000", "26", "13%", "~20 000", "10.0%")
Dataframe3[7,] <- c("Mammals", "336", "57", "17%", "6 399", "5.3%")
Dataframe3[8,] <- c("Octopus, squids", "195", "unknown", "unknown", "800", "24.3%")

table3 <- gt(Dataframe3, rowname_col = "row", groupname_col = "group")

table3 %>%

  tab_style(style = cell_text(weight = "bold"),
    locations = cells_column_labels(c("Taxonomic group", "Species", "Endemics", "Endemics as % of total SA species", "Global estimates of number of species", "Species in SA as % of global total"))
    )  %>%
  
  # Center-align the headings of all columns except the first one
  tab_style(
    style = cell_text(align = "center"),  # Center-align selected column labels
    locations = cells_column_labels(
      columns = c("Species", "Endemics", "Endemics as % of total SA species", 
                  "Global estimates of number of species", "Species in SA as % of global total"))
    )%>%

  # changing font size and title position
   tab_options(
    #heading.align = 'center',
    table.width = pct(80),
    table.background.color = "#FFFFFF00", 
    table.font.color = "black",
    table.font.size = px(13L),
    table.border.top.width = 4,
    data_row.padding = px(10), 
    heading.title.font.size = px(12),
    heading.padding = px(5)
    ) %>%

  cols_width(
    "Taxonomic group" ~ px(100),
    everything() ~ px(15)
  ) %>%
  
  # Left-align the text in all columns
  cols_align(
    align = "center", 
    columns = c("Species", "Endemics", "Endemics as % of total SA species", 
                "Global estimates of number of species", "Species in SA as % of global total")
    #columns = everything()  # Applies to all columns
  )

```

# Biodiversity benefits

Add text

## Benefits of biodiversity in this realm?

### Terrestrial realm

Add text

### Freshwater (inland aquatic) realm

Add text

### Estuarine realm

Add text

### Marine realm

Add text

### Coast cross-cutting realm

Add text
